Miniproject - Elliptic Curves

### 2022-03-31 Elliptic Curves/submission.tex
Removed unused lemma

    Furthermore $E'$ is given by an equation of the form
    \begin{equation*}
        E': y^2 + a_1 x y + a_3 y = x^3 + a_2 x^2
    \end{equation*}
    and $[2](0, 0) = (-a_2, a_1 a_2 - a_3)$.
\end{corollary}
\begin{proof}
    After a translation by $-P$, we may assume that
    \begin{equation*}
        E: y^2 + a'_1 x y + a'_3 y = x^3 + a'_2 x^2 + a'_4 x
    \end{equation*}
    and $P = (0, 0)$.
    Now observe that if $a'_3 = 0$, the line $x = 0$ through $(0, 0)$ and $\O$ meets $E$ in $(0, 0)$ with multiplicity 2, and so $(0, 0) + \O = (0, 0)$, contradicting the assumption that $P$ is not a 2-torsion point.
    Thus $a'_3 \neq 0$ and a shearing with $r = a'_4 / a'_3$ maps $E$ to
    \begin{equation*}
        E': y^2 + a_1 x y + a_3 y = x^3 + a_2 x^2
    \end{equation*}
    Note that the tangent at $(0, 0)$ now has slope $0$, i.e. is given by $y = 0$.
    Furthermore, the third point of intersection of the tangent and $E$ is $(-a_2, 0)$.
    The line through $\O$ and $(-a_2, 0)$ is given by $x = -a_2$ and its third point of intersection with $E$ is then $(-a_2, a_1a_2 - a_3)$. 
\end{proof}
I came up with the next lemma to make my first proof of 3(i) work.
Since then, I have found a simpler proof that does not require the lemma anymore, but I found it beautiful and did not want to delete it.
\begin{lemma}
    \label{prop:share_subgroup_equal}
    Let $E, E'$ be Elliptic Curves defined over any field $k$, and assume they share a cyclic subgroup of order $n \geq 5$\footnote{Technically, we can also allow infinite order here.}.
    With this, we mean there is a point $P \in E \cap E'$ of order $n$ such that
    \begin{equation*}
        G := \langle P \rangle_E \subseteq E' \quad \text{and} \quad \restr{+_E}{G \times G} = \restr{+_{E'}}{G \times G}
    \end{equation*}
    Then $E = E'$ (in the sense that they have the same defining equation).
\end{lemma}
\begin{proof}
    Consider some point $[i] P = (a, b) \in E \cap E'$.
    With $P' = (a', b') := -[2i] P \in E \cap E'$ we see that $P + P + P' = \O$ and so $P, P, P'$ are colinear\footnote{We mean that the line through $P$ and $P'$ meets $E$ resp. $E'$ at $P$ with multiplicity 2.}.
    In particular, the tangent on $E$ resp. on $E'$ at $P \in E \cap E'$ both have the slope $(b - b')/(a - a')$\footnote{Or infinity if $a = a'$, but importantly, the slopes are equal.}.

    Since $P$ has order at least 5, observe that $P, [2]P, [3]P, [4]P$ and $[5]P$ are all different.
    Furthermore, since $E$ and $E'$ have the same tangent slope at each $[i] P$, note that $E$ meets $E'$ in $[i] P$ with multiplicity 2.
    So $E$ meets $E'$ in at least 10 points (counting multiplicity), which is greater than the product of their degrees $9 = 3 \cdot 3$.
    By Bezout's theorem \cite[Corollary I.7.8]{hartshorne}, it follows that $E$ and $E'$ share an irreducible component of dimension $\geq 1$, but since both are Elliptic Curves, they are irreducible of dimension $1$, so we find $E = E'$.
\end{proof}
Finally we can proof the statement on 5-torsion points.
\begin{prop}[3(i)]
    Let $E$ be an Elliptic Curve defined over $\Q$.
    Then $E(\Q)$ has a nontrivial 5-torsion point if and only if $E$ is isomorphic (over $\Q$) to an Elliptic Curve given by an equation of the form
    \begin{equation*}
        y^2 + (v + 1)x y + v y = x^3 + v x^2
    \end{equation*}
\end{prop}
\begin{proof}
    By Corollary~\ref{prop:point_normalization}, we can assume that $E$ is given as
    \begin{equation*}
        E: y^2 + a_1 x y + a_3 y = x^3 + a_2 x^2
    \end{equation*}
    and $(0, 0)$ is a 5-torsion point of $E$ such that the tangent at $(0, 0)$ on $E$ is given by $y = 0$.
    Further, have $[2](0, 0) = (-a_2, a_1 a_2 - a_3)$.
    After applying a scaling, assume further that $a_2 = a_3$.
    Now define $\beta = a_2(a_1 - 1)$.
    Thus $[2](0, 0) = (-a_2, \beta)$.
    By computing the third point of intersection between $E$ and the lines $x = 0$ resp. $x = -a_2$ we see that 
    \begin{equation*}
        [3](0, 0) = -[2](0, 0) = (-a_2, 0), \quad [4](0, 0) = (0, -a_3)
    \end{equation*}
    Now consider the tangent at $[2](0, 0) = (-a_2, \beta)$.
    It has slope
    \begin{align*}
        \lambda =& \frac {3 a_2^2 - 2 a_2^2 - a_1 \beta} {2\beta - a_1 a_2 + a_3} = \frac {a_2^2 - a_1\beta} \beta = \frac {a_2^2 - a_1a_2(a_1 - 1)} \beta = \frac {(a_2 - a_1(a_1 - 1))a_2} {a_2(a_1 - 1)} \\
        =& \frac {a_2 - a_1(a_1 - 1)} {a_1 - 1} = \frac {a_2} {a_1 - 1} - a_1
    \end{align*}
    Since $[4](-a_2, \beta) = [4](0, 0) = -(0, 0)$, observe that $(0, 0)$ must be a point on the tangent $y = \lambda (x + a_2) + \beta$.
    Thus $\lambda a_2 + \beta = 0$ and so
    \begin{equation*}
        \frac {a_2^2} {a_1 - 1} - a_1 a_2 + a_2(a_1 - 1) = 0
    \end{equation*}
    Clearly $a_2 \neq 0$ and thus
    \begin{equation*}
        0 = a_2 - a_1 (a_1 - 1) + (a_1 - 1)^2 = a_2 - a_1 - 1
    \end{equation*}
    So $a_1 = a_2 + 1$ and the claim follows with $v = a_2 = a_3$.
\end{proof}
More or less the same approach works for other $n \neq 5$.
However, the case of 2-torsion points is special (but nevertheless quite easy).
The following is IMO the nicest form for curves with a nontrivial 2-torsion point, even though there are still 2 parameters (and not just one).
\begin{prop}[3(ii) - $2$-torsion points]
    Let $E$ be an Elliptic Curve defined over $\Q$.
    Then $E(\Q)$ has a nontrivial 2-torsion point if and only if $E$ is isomorphic (over $\Q$) to an Elliptic Curve given by an equation of the form
    \begin{equation*}
        y^2 = x^3 + a_2 x^2 + a_4 x
    \end{equation*}
\end{prop}
\begin{proof}
    Assume $E: y^2 = x^3 + a_2 x^2 + a_4 x$ is an Elliptic Curve.
    Then clearly $(0, 0) \in E(\Q)$ and $(0, 0) = -(0, 0)$, so $(0, 0)$ is nontrivial 2-torsion point.
    Hence, there is a nontrivial 2-torsion point in $E'(\Q)$ for all Elliptic Curves $E'$ that are isomorphic (over $\Q$) to $E$.

>>>>

    Furthermore $E'$ is given by an equation of the form
    \begin{equation*}
        E': y^2 + a_1 x y + a_3 y = x^3 + a_2 x^2
    \end{equation*}
    and $[2](0, 0) = (-a_2, a_1 a_2 - a_3)$.
\end{corollary}
\begin{proof}
    After a translation by $-P$, we may assume that
    \begin{equation*}
        E: y^2 + a'_1 x y + a'_3 y = x^3 + a'_2 x^2 + a'_4 x
    \end{equation*}
    and $P = (0, 0)$.
    Now observe that if $a'_3 = 0$, the line $x = 0$ through $(0, 0)$ and $\O$ meets $E$ in $(0, 0)$ with multiplicity 2, and so $(0, 0) + \O = (0, 0)$, contradicting the assumption that $P$ is not a 2-torsion point.
    Thus $a'_3 \neq 0$ and a shearing with $r = a'_4 / a'_3$ maps $E$ to
    \begin{equation*}
        E': y^2 + a_1 x y + a_3 y = x^3 + a_2 x^2
    \end{equation*}
    Note that the tangent at $(0, 0)$ now has slope $0$, i.e. is given by $y = 0$.
    Furthermore, the third point of intersection of the tangent and $E$ is $(-a_2, 0)$.
    The line through $\O$ and $(-a_2, 0)$ is given by $x = -a_2$ and its third point of intersection with $E$ is then $(-a_2, a_1a_2 - a_3)$. 
\end{proof}
Now we can proof the statement on 5-torsion points.
\begin{prop}[3(i)]
    Let $E$ be an Elliptic Curve defined over $\Q$.
    Then $E(\Q)$ has a nontrivial 5-torsion point if and only if $E$ is isomorphic (over $\Q$) to an Elliptic Curve given by an equation of the form
    \begin{equation*}
        y^2 + (v + 1)x y + v y = x^3 + v x^2
    \end{equation*}
\end{prop}
\begin{proof}
    By Corollary~\ref{prop:point_normalization}, we can assume that $E$ is given as
    \begin{equation*}
        E: y^2 + a_1 x y + a_3 y = x^3 + a_2 x^2
    \end{equation*}
    and $(0, 0)$ is a 5-torsion point of $E$ such that the tangent at $(0, 0)$ on $E$ is given by $y = 0$.
    Further, have $[2](0, 0) = (-a_2, a_1 a_2 - a_3)$.
    After applying a scaling, assume further that $a_2 = a_3$.
    Now define $\beta = a_2(a_1 - 1)$.
    Thus $[2](0, 0) = (-a_2, \beta)$.
    By computing the third point of intersection between $E$ and the lines $x = 0$ resp. $x = -a_2$ we see that 
    \begin{equation*}
        [3](0, 0) = -[2](0, 0) = (-a_2, 0), \quad [4](0, 0) = (0, -a_3)
    \end{equation*}
    Now consider the tangent at $[2](0, 0) = (-a_2, \beta)$.
    It has slope
    \begin{align*}
        \lambda =& \frac {3 a_2^2 - 2 a_2^2 - a_1 \beta} {2\beta - a_1 a_2 + a_3} = \frac {a_2^2 - a_1\beta} \beta = \frac {a_2^2 - a_1a_2(a_1 - 1)} \beta = \frac {(a_2 - a_1(a_1 - 1))a_2} {a_2(a_1 - 1)} \\
        =& \frac {a_2 - a_1(a_1 - 1)} {a_1 - 1} = \frac {a_2} {a_1 - 1} - a_1
    \end{align*}
    Since $[4](-a_2, \beta) = [4](0, 0) = -(0, 0)$, observe that $(0, 0)$ must be a point on the tangent $y = \lambda (x + a_2) + \beta$.
    Thus $\lambda a_2 + \beta = 0$ and so
    \begin{equation*}
        \frac {a_2^2} {a_1 - 1} - a_1 a_2 + a_2(a_1 - 1) = 0
    \end{equation*}
    Clearly $a_2 \neq 0$ and thus
    \begin{equation*}
        0 = a_2 - a_1 (a_1 - 1) + (a_1 - 1)^2 = a_2 - a_1 - 1
    \end{equation*}
    So $a_1 = a_2 + 1$ and the claim follows with $v = a_2 = a_3$.
\end{proof}
More or less the same approach works for other $n \neq 5$.
However, the case of 2-torsion points is special (but nevertheless quite easy).
The following is IMO the nicest form for curves with a nontrivial 2-torsion point, even though there are still 2 parameters (and not just one).
\begin{prop}[3(ii) - $2$-torsion points]
    Let $E$ be an Elliptic Curve defined over $\Q$.
    Then $E(\Q)$ has a nontrivial 2-torsion point if and only if $E$ is isomorphic (over $\Q$) to an Elliptic Curve given by an equation of the form
    \begin{equation*}
        y^2 = x^3 + a_2 x^2 + a_4 x
    \end{equation*}
\end{prop}
\begin{proof}
    Assume $E: y^2 = x^3 + a_2 x^2 + a_4 x$ is an Elliptic Curve.
    Then clearly $(0, 0) \in E(\Q)$ and $(0, 0) = -(0, 0)$, so $(0, 0)$ is nontrivial 2-torsion point.
    Hence, there is a nontrivial 2-torsion point in $E'(\Q)$ for all Elliptic Curves $E'$ that are isomorphic (over $\Q$) to $E$.
